---
title: 'University_Debt: Small V Private'
author: "Khawaja Hussain Ahmed"
date: "2023-10-19"
---

<center><h2>  __Final Report__  </h2> </center>
\
\

The purpose of the study is to validate if there is a statistically significant difference between the average debt of graduates of small,private universities and graduates of large, private universities.

\
\

We use two independent samples hypothesis testing procedures to answer this question. To conduct our hypothesis test, we proceed as follows:

\

Suppose that $μ_small$ is the average debt of graduates from small private universities and $μ_large$ is the average debt of graduates from large, private universities. If there is a significant difference, then $μ_small-μ_large>0$. If there isn't a statistical difference, then $μ_1-μ_2=0$. So, we set up a null hypothesis $H_0: μ_small-μ_large=0$ and an alternative hypothesis $H_a: μ_small-μ_large>0$.

\

We will use two independent sample t-test for hypothesis testing. We assume that two samples:(a) average debt for graduates of small, private universities; and (b) average debt for graduates of large, private universities were randomly taken from two underlying independent populations. To conduct our test, we need to check that these two samples were collected from two normally distributed populations and the (population) standard deviations of average debt for small,private universities and large,private universities are equal but unknown. We draw a side-by-side comparison and plot two box-plots of these two samples, to find that the sample evidences approximately support the normality assumptions.

```{r, results="hide",echo=FALSE}

college = read.csv("college_debt.csv")
college$Debt = as.integer(gsub("[\\$,]", "", college$Debt))

```


```{r, results='hide',echo=FALSE}
library(ggplot2)

ggplot(data=college, aes(x=Debt, y="", fill=Status)) + 
  geom_boxplot()+
  facet_wrap(~Status)+
  
  theme(plot.title = element_text(vjust =-100, hjust=.5),
        panel.spacing = unit(1, "cm"),
        legend.position = "none",
        axis.title.x=element_text(hjust=0.5),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.margin=margin(t=2,r=0,b=2,l=2,unit="cm"))+
  ylab("")+
  xlab(" Average Debt")+
  ggtitle("Figure 1. Boxplot of Average Debt for Small and Large Private Universities \n") 
```
\
In addition, we produce quantile-quantile (Q-Q) plot for each sample to check whether the two data sets come from populations with a common distribution. The Q-Q plots also support the normality assumption about the populations.
\

```{r, results="hide",echo=FALSE}
ggplot(data = college, aes(sample=Debt)) +
  geom_qq() + 
  geom_qq_line(color="red") + 
  facet_wrap( ~ Status)+
 
  theme(plot.title = element_text(vjust =-100, hjust=.5),
        panel.spacing = unit(1, "cm"),
        legend.position = "none",
        axis.title.x=element_text(hjust=0.5),
        axis.title.y=element_text(angle=0),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        plot.margin=margin(t=2,r=-1,b=2,l=0,unit="cm"))+
    ggtitle("Figure 2. Q-Q plots of the sample data\n")+
    xlab("Normal theoritical quantiles")+
    ylab("Sample data\n quantiles")


```

\

```{r, results='hide',echo=FALSE}
tapply(college$Debt,college$Status, sd)
```

To check the equality variances, we calculate  $s_small/s_large =5453.599/7034.909=0.78$`, which lies between 0.5 and 2, so this assumption is also satisfied.

\


```{r,results='hide',echo=FALSE}
t.test(college$Debt~college$Status,
       alternative=c("greater"),
       mu=0,paired=F,var.equal=T)

```


The statistical test results found that $t=-2.2974$ and $P-value= 0.9879$, that is we do not reject $H_0$ and this implies that there is no statistically significant difference.

\

```{r, results="hide",echo=FALSE}
tapply(college$Debt,college$Status, mean)
```

The average of small,private universities is $\bar{x}=32239.26$ and the average of large,private universities is $\bar{x}=29124$.

\

In summary, the t-test results suggest that there is not a statistically significant difference between the means of the groups being compared, as indicated by the high p-value.